Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:
1. https://classroom.udacity.com/courses/ud651/lessons/755618712/concepts/8140986010923
1. http://hci.stanford.edu/publications/2013/invisibleaudience/invisibleaudience.pdf

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv("pseudo_facebook.tsv", sep="\t")

qplot(x = age, y = friend_count, data = pf)

```

***

#### What are some things that you notice right away?
Response:
That the distribution of the data isn't normal.
It appears that a lot of data are on the left and right of the chart.
And there's also a spike of friends count for users who is around 70 years old.
***

### ggplot Syntax
Notes:
1. http://ggplot2.tidyverse.org/reference/
1. http://bbs.ceb-institute.org/wp-content/uploads/2011/09/handout_ggplot2.pdf

```{r ggplot Syntax}
ggplot(data =pf, aes(x= age, y= friend_count)) +
    geom_point() + 
    scale_x_continuous(breaks=seq(13,90, 2), limits = c(13,90))

summary(pf$age)

```

***

### Overplotting
Notes:
1. http://ggplot2.tidyverse.org/reference/geom_point.html

```{r Overplotting}
# it will take twenty friend count to get a dark dot
ggplot(data = pf, aes(x=age, y=friend_count)) + 
    geom_jitter(alpha=1/20) +
    xlim(13, 90)

```

#### What do you notice in the plot?
Response:
That the data is a bit more spread out now.
There's still a high concentration of  younger users with high friends  count which seems normal.
The users that are roughly 70 years old have  have a common friends count of less than 1500. And those that have a friend count of more than 1,500 isn't that common.
***

### Coord_trans()
Notes:
1. http://ggplot2.tidyverse.org/reference/coord_trans.html

```{r Coord_trans()}

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(data = pf, aes(x=age, y=friend_count)) + 
    geom_jitter(alpha=1/20, position=position_jitter(h = 0)) +
    xlim(13, 90) +
    coord_trans(y="sqrt")

ggplot(data = pf, aes(x=age, y=friend_count)) + 
    geom_point(alpha=1/20) +
    xlim(13, 90) +
    coord_trans(y="sqrt")

```

#### What do you notice?
That the values are mo spread out for users below the age of 25.
By sqrt the y, it makes the users with a friend count of below 1000 easier to spot.
***

### Alpha and Jitter
Notes:
1. http://ggplot2.tidyverse.org/reference/position_jitter.html

```{r Alpha and Jitter}
# This programming assignment
# will not be graded, but when you
# submit your code, the assignment
# will be marked as correct. By submitting
# your code, we can add to the feedback
# messages and address common mistakes
# in the Instructor Notes.

# You can assess your work by watching
# the solution video.


# Examine the relationship between
# friendships_initiated (y) and age (x)
# using the ggplot syntax.
ggplot(data= pf, aes(x =age, y=friendships_initiated)) +
    geom_point(show.legend = T) 

# We recommend creating a basic scatter
# plot first to see what the distribution looks like.
# and then adjusting it by adding one layer at a time.
ggplot(data= pf, aes(x =age, y=friendships_initiated)) +
    geom_jitter(show.legend = T, alpha = 1/20) +
    scale_x_continuous(breaks=seq(13,90, 2), limits = c(13,90))

ggplot(data= pf, aes(x =age, y=friendships_initiated)) +
    geom_jitter(show.legend = T, alpha = 1/20, position = position_jitter(h=0)) +
    scale_x_continuous(breaks=seq(13,90, 2), limits = c(13,90)) +
    coord_trans(y = "sqrt")

# What are your observations about your final plot?
#1. Resemble the friend count's plot, which we see a high number of users with friends count which are below the age of 25.
#1. We also see the similar spike for the user around the age of 68.
#1. There appears to be more users who are above the age of 73 that doesn't have low friendship initiated count.

# Remember to make adjustments to the breaks
# of the x-axis and to use apply alpha and jitter.

# ENTER ALL OF YOUR CODE FOR YOUR PLOT BELOW THIS LINE.
# =======================================================

```

## plot analysis
1. Resemble the friend count's plot, which we see a high number of users with friends count which are below the age of 25.
1. We also see the similar spike for the user around the age of 68.
1. There appears to be more users who are above the age of 73 that doesn't have low friendship initiated count.

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:
1. https://blog.rstudio.com/2014/01/17/introducing-dplyr/
1. https://www.r-bloggers.com/hadley-wickham-presents-dplyr-at-user-2014/
1. https://www.r-bloggers.com/hadley-wickhams-dplyr-tutorial-at-user-2014-part-1/
1. https://www.r-bloggers.com/hadley-wickhams-dplyr-tutorial-at-user-2014-part-2/
1. https://rollingyours.wordpress.com/2014/10/20/the-lapply-command-101/
1. http://ggplot2.tidyverse.org/reference/geom_path.html

```{r Conditional Means}
library(dplyr)

age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups, 
        fc_mean = mean(friend_count),
        fc_median = median(friend_count),
        n = n())
pf.fc_by_age <- arrange(pf.fc_by_age, age)

## an alternative way to get this done
pf.fc_by_age <- pf %>%
    group_by(age) %>%
    summarise(fc_mean = mean(friend_count),
              fc_median = median(friend_count),
              n = n()) %>%
    arrange(age)


head(pf.fc_by_age)
```

Create your plot!

```{r Conditional Means Plot}
# This programming assignment
# will not be graded, but when you
# submit your code, the assignment
# will be marked as correct. By submitting
# your code, we can add to the feedback
# messages and address common mistakes
# in the Instructor Notes.

# You can assess your work by watching
# the solution video.


# Plot mean friend count vs. age using a line graph.
# Be sure you use the correct variable names
# and the correct data frame. You should be working
# with the new data frame created from the dplyr
# functions. The data frame is called 'pf.fc_by_age'.

# Use geom_line() rather than geom_point to create
# the plot. You can look up the documentation for
# geom_line() to see what it does.

# ENTER ALL OF YOUR CODE TO CREATE THE PLOT BELOW THIS LINE.
# ===========================================================
ggplot(data = pf.fc_by_age, aes(x=age, y=fc_mean)) +
    geom_line() +
    scale_x_continuous(breaks=seq(13, 120, 3), limits = c(13,120))

```

***

### Overlaying Summaries with Raw Data
Notes:
1. http://www.r-tutor.com/elementary-statistics/numerical-measures/percentile

```{r Overlaying Summaries with Raw Data}

#scale_y_continuous is used for subsetting the data with the help of xlim and ylim
#coord_cartesian is used to zoom in on a plot programatically

#ggplot(data = pf, aes(x=age, y=friend_count)) +
#    coord_cartesian(xlim(13, 90), ylim=c(0,1000)) +
#    geom_point(alpha=1/20)
    #+ geom_line(stat="summary", fun.y = mean)
    #+ geom_line(stat="summary", fun.y= quantile, fun.args = list(probs=.1), linetype=2, color="orange")
    #+ geom_line(stat="summary", fun.y= quantile, fun.args = list(probs=.9), linetype=2, color="orange")
    #+ geom_line(stat="summary", fun.y= quantile, fun.args = list(probs=.5), linetype=2, color="orange")

ggplot(data = pf, aes(x=age, y=friend_count)) + 
    scale_x_continuous(limit = c(13, 70)) + 
    scale_y_continuous(limit = c(0,1000)) +
    coord_trans(y="sqrt") +
    geom_point(alpha=1/20, position=position_jitter(h = 0), color="blue") +
    geom_line(stat="summary", fun.y = mean) +
    geom_line(stat="summary", fun.y= quantile, fun.args = list(probs=.1), linetype=2, color="orange") +
    geom_line(stat="summary", fun.y= quantile, fun.args = list(probs=.5), linetype=2, color="orange") +
    geom_line(stat="summary", fun.y= quantile, fun.args = list(probs=.9), linetype=2, color="orange")
ggsave(filename="overlayRawData.png", width = 15, height = 4, dpi = 75, units = "in", device='png')

```

```{r}
ggplot(data = pf, aes(x=age, y=friend_count)) + 
    coord_cartesian(xlim=c(13,70), ylim=c(0,1000)) +
    geom_point(alpha=1/20, position=position_jitter(h = 0), color="blue") +
    geom_line(stat="summary", fun.y = mean) +
    geom_line(stat="summary", fun.y= quantile, fun.args = list(probs=.1), linetype=2, color="orange") +
    geom_line(stat="summary", fun.y= quantile, fun.args = list(probs=.5), linetype=2, color="orange") +
    geom_line(stat="summary", fun.y= quantile, fun.args = list(probs=.9), linetype=2, color="orange")
```

#### What are some of your observations of the plot?
Response:
1. Most users have less than 250 friends.
1. The upper fifty percent have a wider range of friends. That explains why the median is higher than the 50% quintile value.


***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:
1. the histogram that she created wasn't really intuitive on first view and require additional explanation to make it clear what she was comparing against.

***

### Correlation
Notes:
1. https://s3.amazonaws.com/udacity-hosted-downloads/ud651/correlation_images.jpeg
1. http://www.r-tutor.com/elementary-statistics/numerical-measures/correlation-coefficient
1. https://classroom.udacity.com/courses/ud201/lessons/1345848540/concepts/1715827370923

```{r Correlation}
cor.test(pf$age, pf$friend_count, method='pearson')

round(cor(pf$age, pf$friend_count),3)

with(pf, cor.test(age, friend_count, method='pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

