Lesson 3
========================================================

### References
1. https://flowingdata.com/2014/02/27/how-to-read-histograms-and-use-them-in-r/
1. https://github.com/jrnold/ggthemes#examples
1. http://www.cookbook-r.com/Graphs/Facets_(ggplot2)/
1. https://en.wikipedia.org/wiki/Top-coded
1. http://ggplot2.tidyverse.org/reference/scale_continuous.html
1. https://www.statmethods.net/input/missingdata.html
1. http://ggplot2.tidyverse.org/reference/theme.html
1. https://s3.amazonaws.com/udacity-hosted-downloads/ud651/pseudo_facebook.tsv

***

### What to Do First?
Notes:
Create a histogram of the data for the initial analysis.
***

```{r}
getwd()
list.files()
pf <- read.csv("pseudo_facebook.tsv", sep='\t')
names(pf)

```


### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
#install.packages('ggthemes', dependencies = TRUE)
library(ggthemes)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)

ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31)

```

***

#### What are some things that you notice about this histogram?
Response:
The 1st day of the month have an unusual higher number of user counts, whereas the last day of the month has a lower number of users count due to the fact not every month has a 31st.
***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:
20 views from friends

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:
5%
***

### Perceived Audience Size
Notes:


***
### Faceting
Notes:

```{r Faceting}
ggplot(data = pf, aes(x = dob_day)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month, ncol = 3)
```

#### Let???s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira???s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram()

qplot(data=pf, x= friend_count)
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:


```{r Limiting the Axes}
qplot(data=pf, x=friend_count, xlim= c(0,1000))

qplot(data=pf, x=friend_count) +
    scale_x_continuous(limits = c(0,1000))

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000))
```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender, ncol=2)

qplot(x = friend_count, data = pf) +
  facet_grid(gender ~ .)

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf,!is.na(pf$gender))) +
  facet_wrap(~gender)

ggplot(aes(x = friend_count), data = subset(pf,!is.na(pf$gender))) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)

by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response: women

#### What's the difference between the median friend count for women and men?
Response: 22

#### Why would the median be a better measure than the mean?
Response: not affected by the outlier value which in this dataset is a very larger number.

***

### Tenure
Notes:

```{r Tenure}
qplot(data=pf, x = tenure, color=I('black'), fill=I('#099DD9'), binwidth=30)

ggplot(aes(x = tenure), data = pf) +
  geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(data=pf, x = tenure/365, color=I('black'), fill=I('#099DD9'), binwidth=.25) +
 scale_x_continuous(breaks=seq(1,7,1), limits=c(0, 7))
 
ggplot(aes(x = tenure/365), data = pf) +
  geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(data=pf, x = tenure/365, color=I('black'), fill=I('#099DD9'), binwidth=.25,
      xlab="Number of years using Facebook", ylab="Number of users in sample") +
 scale_x_continuous(breaks=seq(1,7,1), limits=c(0, 7))
```

***

### User Ages
Notes:

```{r User Ages}
qplot(data=pf, x=age, color=I('black'), fill=I('green'), binwidth=1, 
      xlab="Age of facebook users", ylab="Number of users in sample") +
    scale_x_continuous(breaks=seq(0,115, 5), limits=c(13,113))

summary(pf$age)
```

#### What do you notice?
Response:
There are no users that is younger than 15.
***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:
1. https://www.r-statistics.com/2013/05/log-transformations-for-skewed-and-wide-distributions-from-practical-data-science-with-r/
1. https://en.wikipedia.org/wiki/Normal_distribution
1. https://en.wikipedia.org/wiki/Linear_regression#Assumptions
1. http://lightonphiri.org/blog/ggplot2-multiple-plots-in-one-graph-using-gridextra
1. https://www.r-statistics.com/2013/05/log-transformations-for-skewed-and-wide-distributions-from-practical-data-science-with-r/


```{r Add a Scaling Layer}
p1 = qplot(data=pf, x=pf$friend_count, color=I('black'), fill=I('#099DD9'),
      xlab="Number of friends per Facebook user", ylab="Number of users in sample")

summary(pf$friend_count)

summary (log10(pf$friend_count + 1))

summary (sqrt(pf$friend_count))

install.packages('gridExtra')
library(gridExtra)

p2 = qplot(data=pf, x=log10(pf$friend_count), color=I('black'), fill=I('orange'),
      xlab="Number of friends per Facebook user in log10", ylab="Number of users in sample")

p3 = qplot(data=pf, x=sqrt(pf$friend_count), color=I('black'), fill=I('green'),
      xlab="Number of friends per Facebook user in sqrt", ylab="Number of users in sample")

grid.arrange(p1,p2,p3, ncol=2)
```

``` {r}
p1 = ggplot(data=pf, aes(x=friend_count)) + geom_histogram(color=I('black'), fill=I('#099DD9'))

p2 = p1 + scale_x_log10()

p3 = p1 + scale_x_sqrt()

grid.arrange(p1,p2,p3, ncol=2)
```

```{r}
logScale <- qplot(data=pf, x=log10(friend_count))
countScale <- ggplot(aes(x=friend_count), data=pf) + geom_histogram() + scale_x_log10()
logScale2 <- qplot(data=pf, x=friend_count) + scale_x_log10()
grid.arrange(logScale, countScale, logScale2)
```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(x = friend_count, data = subset(pf,!is.na(pf$gender))) +
  facet_wrap(~gender)

qplot(x = friend_count, y = ..count../sum(..count..),
      data = subset(pf,!is.na(pf$gender)), binwidth=10, color=gender, geom='freqpoly',
      xlab="Friend count", ylab="Proportion of users with that friend count") +
    scale_x_continuous(limits = c(0,1000), breaks= seq(0,1000,50))

ggplot(aes(x = friend_count, y = ..count../sum(..count..)),
       data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth=5) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 20)) +
  xlab('Friend Count') +
  ylab('Proportion of users with that friend count')
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
qplot(x = www_likes,
      data = subset(pf,!is.na(pf$gender)), color=gender, geom='freqpoly',
      xlab="Likes count", ylab="Proportion of users with that friend count") +
    scale_x_log10()

summary(pf$www_likes)

by(pf$www_likes, pf$gender, sum)

```


***

### Box Plots
Notes:

1. https://flowingdata.com/2008/02/15/how-to-read-and-use-a-box-and-whisker-plot/
1. https://en.wikipedia.org/wiki/Interquartile_range
1. https://en.wikipedia.org/wiki/File:Boxplot_vs_PDF.svg

```{r Box Plots}
qplot(data= subset(pf,!is.na(gender) & friend_count < 1000),
                   x=gender, y=friend_count, geom='boxplot') +
  
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
box1 <- qplot(data= subset(pf,!is.na(gender) & friend_count < 1000),
                   x=gender, y=friend_count, geom='boxplot')

box2 <- qplot(data= subset(pf,!is.na(gender)),
                   x=gender, y=friend_count, geom='boxplot', ylim=c(0,1000))

box3 <- qplot(data= subset(pf,!is.na(gender)),
                   x=gender, y=friend_count, geom='boxplot') +
    scale_y_continuous(limits = c(0,1000))

grid.arrange(box1, box2, box3)
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(data= subset(pf,!is.na(gender)),
                   x=gender, y=friend_count, geom='boxplot') +
    coord_cartesian(ylim = c(0,250))

by(pf$friend_count, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
names(pf)

qplot(data= subset(pf,!is.na(gender)),
                   x=gender, y=friendships_initiated, geom='boxplot') +
    coord_cartesian(ylim = c(0,150))

by(pf$friendships_initiated, pf$gender, summary)
```

Response:

***
I did a boxplot comparison against gender and friendship initiated to get a first overview of the data spread between the two groups. It does appears that female group has a slightly median value. Thus to zoom in on the exact value, I use the by function to get the summary value of the dataset. And indeed the female group has a higher median value of 49 versus the male group which has a median value of 44. Comparing against the mean value as well, that proves to be consistent with the female group has a higher average value at 113.9 versus the male group with a mean value of 103.1

### Getting Logical
Notes:

```{r Getting Logical}
mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary (pf$mobile_check_in)

check_in_pct <-  round(sum(pf$mobile_check_in==1)/length(pf$mobile_check_in) * 100,0)
print (check_in_pct)
```

Response:

***

### Analyzing One Variable
Reflection:

***

Notes:

distribution, missing values, outliers
boxplots, histogram
changing the data using log10, sqrt may uncover additional patterns.

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!