---
title: "Lesson 4 exercise"
output: html_notebook
---

```{r diamonds}
library(ggplot2) #must load the ggplot package first
data(diamonds) #loads the diamonds data set since it comes with the ggplot package

summary(diamonds)

?diamonds

summary(diamonds$color)

str(diamonds)
```

```{r}
ggplot(data=diamonds, aes(x=price, y=x)) +
  geom_point()

```

## observation
There appears to be a positive correlation between price and x


```{r}

round(with(diamonds, cor(price, x)),2)
round(with(diamonds, cor(price, y)),2)
round(with(diamonds, cor(price, z)),2)

range(diamonds$price)
```
```{r}
ggplot(data=diamonds, aes(x=price, y=depth)) +
  geom_point()

```
```{r}
# Change the code to make the transparency of the
# points to be 1/100 of what they are now and mark
# the x-axis every 2 units. See the instructor notes
# for two hints.
ggplot(data=diamonds, aes(x=price, y=depth)) +
  geom_jitter(alpha = 1/100) +
  scale_x_continuous(breaks=seq(0,18823,500))

```
```{r}

round(with(diamonds, cor(price, depth)),2)
```



```{r}
# Create a scatterplot of price vs carat
# and omit the top 1% of price and carat
# values.
ggplot(data=diamonds, aes(x=carat, y=price)) +
  geom_point() +
  xlim(0, quantile(diamonds$carat, 0.99)) +
  ylim(0, quantile(diamonds$price, 0.99))
```
```{r}
# Create a scatterplot of price vs. volume (x * y * z).
# This is a very rough approximation for a diamond's volume.
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

# Create a new variable for volume in the diamonds data frame.
# This will be useful in a later exercise.
ggplot(data=diamonds, aes(x=volume, y=price)) +
  geom_point() 
```
## Observation
Outliers are present.
There appears to be a good correlation between the two variable.
Would need to deep dive in after excluding the outliers.

```{r}
library(plyr)
count(diamonds$volume == 0)
detach("package:plyr", unload=TRUE)

```

```{r}
round(with(subset(diamonds, volume > 0 & volume < 800), cor(price, volume)), 2)
```

Notes:
1. types of smoother https://stats.idre.ucla.edu/r/faq/how-can-i-explore-different-smooths-in-ggplot2/

```{r}
# Subset the data to exclude diamonds with a volume
# greater than or equal to 800. Also, exclude diamonds
# with a volume of 0. Adjust the transparency of the
# points and add a linear model to the plot. (See the
# Instructor Notes or look up the documentation of
# geom_smooth() for more details about smoothers.)
ggplot(data=subset(diamonds, volume > 0 & volume < 800), aes(x=price, y=volume)) +
  geom_jitter(alpha= 1/20) +
  geom_smooth()

# We encourage you to think about this next question and
# to post your thoughts in the discussion section.

# Do you think this would be a useful model to estimate
# the price of diamonds? Why or why not?

```
## Observation
Since the plot shows strong correlation especially on the points which are below $10K, it should be a useful model.
Notes: https://discussions.udacity.com/t/lesson-6-adjustments-price-vs-volume/294947/5

```{r}
# Use the function dplyr package
# to create a new data frame containing
# info on diamonds by clarity.

# Name the data frame diamondsByClarity

# The data frame should contain the following
# variables in this order.

#       (1) mean_price
#       (2) median_price
#       (3) min_price
#       (4) max_price
#       (5) n

# where n is the number of diamonds in each
# level of clarity.
library(dplyr)
diamonds_group <- group_by(diamonds, clarity)
diamonds.diamondsByClarify <- summarise(diamonds_group,
                                        mean_price = mean(price),
                                        median_price = median(price),
                                        min_price = min(price),
                                        max_price = max(price),
                                        n = n())

```


```{r}
# We’ve created summary data frames with the mean price
# by clarity and color. You can run the code in R to
# verify what data is in the variables diamonds_mp_by_clarity
# and diamonds_mp_by_color.

# Your task is to write additional code to create two bar plots
# on one output image using the grid.arrange() function from the package
# gridExtra.

diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- ggplot(data=diamonds_mp_by_clarity, aes(x=clarity, y=mean_price)) +
  geom_bar(stat='identity')
p2 <- ggplot(data=diamonds_mp_by_color, aes(x=color,y=mean_price)) +
  geom_bar(stat='identity')

grid.arrange(p1,p2)
```

## Observation
Somehow the most 
Notes: https://discussions.udacity.com/t/trends-in-mean-price/255805/5