---
title: "R Notebook"
output: html_notebook
---

#### References
1. https://flowingdata.com/2015/02/18/loading-data-and-basic-formatting-in-r/
1. https://www.statmethods.net/management/subset.html
1. https://redditblog.com/2011/09/12/who-in-the-world-is-reddit-results-are-in/
1. https://www.statmethods.net/input/datatypes.html
1. https://stats.idre.ucla.edu/r/modules/factor-variables/

#### Shortcuts
1. Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 
1. Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.
1. When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

```{r}
getwd()
setwd('/Users/stevetan/Dropbox/0udacity/data analyst nanodegree/eda/')
statesInfo = read.csv("stateData.csv")
View (statesInfo)
```

```{r}
stateSubset <- subset(statesInfo, murder > 10)
head (stateSubset)
dim (stateSubset)

stateSubsetBracket <- statesInfo[statesInfo$murder > 10,]
head (stateSubsetBracket)
dim (stateSubsetBracket)

```

```{r}
?cars
str(cars)
```

To embed files

```{r}
plot(cars)
```
Testing tables
```{r}
reddit <- read.csv("reddit.csv")
str(reddit)
print ("summary")
summary(reddit)
print ("table")
table(reddit$employment.status)
```
Ordered factors

```{r}
levels(reddit$age.range)

library(ggplot2)
qplot(data=reddit, x= age.range)
```

```{r}
# Order the factor levels in the age.range variable in order to create
# a graph with a natural order. Look up the documentation for
# the factor function or read through the example in the Instructor Notes.

# Once you're ready, try to write the code to order the levels of
# the age.range variable.

# Be sure you modify the variable in the data frame. That is modify reddit$age.range.
# Don't create a new variable.

# The levels of age.range should take on these values...

#    "Under 18", "18-24", "25-34", "35-44", "45-54", "55-64", "65 or Above"

# This exercise is ungraded. You can check your own work by using the Test Run
# button. Your plot will appear there.

# ENTER YOUR CODE BELOW THE LINE.
# ================================================================================


```

